---
title: "Improving Deconvolution Methods in Biology 
        through Open Innovation Competitions: 
        an Application to the Connectivity Map"

author: 
    - Author 1
    - Author 2
    - ... 

abstract: "Report results fo open innovation competition aimed at solving 
           a gene-related deconvolution problem." 

date: 'Last updated: `r format(Sys.Date(), "%b %d, %Y")`'

keywords:
  - biology
  - open innoation competitions
  - crowdsourcing
  - deconvolution
  - gene expressions
  - cell lines

titlepage: false 
linestretch: 1.25
fontfamily: mathpazo
# fontsize: 11pt
# colorlinks: true
# urlcolor: cyan
# citecolor: blue 
geometry: margin=1.25in
# endfloat: true
# bibliography: races.bib
header-includes:
  - \usepackage{bbm}
  - \newcommand\hl{}
  - \newcommand\Fig{Fig. XXXX}
---


TODOS:

- Intro AB
- Methdos TN [DONE]
- Figure Scatter runtime vs accuracy for top10 TN [DONE]
- Add data


```{r sections, echo=FALSE}
childs <- c("intro.rmd", "methods.rmd", "results.rmd", "disc.rmd", "figures.rmd")
```

```{r child=childs}
```

## To read

- [Compound signature detection on LINCS L1000 big data](https://pubs.rsc.org/en/content/getauthorversionpdf/c4mb00677a) used a fuzzy c-means Gaussian Mixture Model (GMM) to process raw L1000 data, showing better performance compared to KNN. This method is described below:

> To deconvolute such overlapped peaks, we assumed that the fluorophore intensities of each analyte type (corresponding to a specific mRNA type) had a Gaussian distribution. The distribution of the mixture of analytes GeneH(i) and GeneL(i) corresponding to the expression levels of GeneH and GeneL, respectively, should be subject to a bimodal Gaussian distribution, with the proportion of 1.25 to 0.75. We initialized the estimations of the two Gaussian distributions using buzzy c-means clustering [11] and estimated the GMM parameters using the Nelder-Mead method [12]. Thus, the overlapped peaks were deconvoluted as the two estimated Gaussian peaks and the expression levels of the two genes sharing the same analyte were extracted. Mathematical details are included in the Supplementary Methods (the GMM model).

  
- [Deconvolution of linear systems by constrained regression and its relationship to the Wiener theory](https://ieeexplore.ieee.org/abstract/document/4044778)

- [Efficient Bayesian-based multiview deconvolution](https://www.nature.com/articles/nmeth.2929)

- [A Bayesian deconvolution strategy for immunoprecipitation-based DNA methylome analysis](https://www.nature.com/articles/nbt1414)

- [Gene expression deconvolution in linear space](https://www.nature.com/articles/nmeth.1830)

- [Cell type–specific gene expression differences in complex tissues](https://www.nature.com/articles/nmeth.1439)


References
==========

