---
title: "Improving Deconvolution Methods in Biology \
        through Open Innovation Competitions: \
        an Application to the Connectivity Map"

author: 
    - Author 1
    - Author 2
    - ... 

abstract: "Report results fo open innovation competition aimed at solving 
           a gene-related deconvolution problem." 

date: 'Last updated: `r format(Sys.Date(), "%b %d, %Y")`'

keywords:
  - biology
  - open innoation competitions
  - crowdsourcing
  - deconvolution
  - gene expressions
  - cell lines

titlepage: false 
linestretch: 1.25
fontfamily: mathpazo
# fontsize: 11pt
# colorlinks: true
# urlcolor: cyan
# citecolor: blue 
geometry: margin=1.25in
# endfloat: true
bibliography: dpeak.bib
biblatexoptions: "sorting=none"
header-includes:
  - \usepackage{bbm}
  - \newcommand\hl{}
  - \newcommand\Fig[1]{{\it Fig. \ref{fig:#1}}}
  - \usepackage{setspace}
  - \onehalfspacing 
---


\color{red}

TODOS:

- Check length requirements for Nat. Met. 
- Intro [AB]
- Methdos TN [DONE]
- Clustering 
- Add disaggregated data

> A Brief Communication is a more concise format used typically to report a significant improvement to a tried-and-tested method, its modification and adaptation to an important original application, or an important new tool or resource of broad interest for the scientific community. This format typically does not exceed 3 printed pages. Brief Communications begin with a brief unreferenced abstract (3 sentences, no more than 70 words), which will appear on Medline. The title is limited to 10 words (or 90 characters). The main text is typically 1,000-1,500 words, including the abstract and contains no headings with the exception of a single heading for Methods to point readers to the online Methods section providing all technical details necessary for the independent reproduction of the methodology. Brief Communications normally have no more than 2 display items, although this may be flexible at the discretion of the editor, provided the page limit is observed. As a guideline, Brief Communications allow up to 20 references, and article titles are omitted from the reference list.

Brief Communications include received/accepted dates. They may be accompanied by supplementary information. Brief Communications are peer reviewed.

\color{black}

```{r sections, echo=FALSE}
childs <- c("intro.rmd", "methods.rmd", "results.rmd", "disc.rmd", "figures.rmd", "supporting.rmd")
```

```{r child=childs}
```

\clearpage 

## To read

- [Compound signature detection on LINCS L1000 big data](https://pubs.rsc.org/en/content/getauthorversionpdf/c4mb00677a) used a fuzzy c-means Gaussian Mixture Model (GMM) to process raw L1000 data, showing better performance compared to KNN. This method is described below:

> To deconvolute such overlapped peaks, we assumed that the fluorophore intensities of each analyte type (corresponding to a specific mRNA type) had a Gaussian distribution. The distribution of the mixture of analytes GeneH(i) and GeneL(i) corresponding to the expression levels of GeneH and GeneL, respectively, should be subject to a bimodal Gaussian distribution, with the proportion of 1.25 to 0.75. We initialized the estimations of the two Gaussian distributions using buzzy c-means clustering [11] and estimated the GMM parameters using the Nelder-Mead method [12]. Thus, the overlapped peaks were deconvoluted as the two estimated Gaussian peaks and the expression levels of the two genes sharing the same analyte were extracted. Mathematical details are included in the Supplementary Methods (the GMM model).

  
- [Deconvolution of linear systems by constrained regression and its relationship to the Wiener theory](https://ieeexplore.ieee.org/abstract/document/4044778)

- [Efficient Bayesian-based multiview deconvolution](https://www.nature.com/articles/nmeth.2929)

- [A Bayesian deconvolution strategy for immunoprecipitation-based DNA methylome analysis](https://www.nature.com/articles/nbt1414)

- [Gene expression deconvolution in linear space](https://www.nature.com/articles/nmeth.1830)

- [Cell type–specific gene expression differences in complex tissues](https://www.nature.com/articles/nmeth.1439)


References
==========

